### 2020-02-05 presentations/jsps_2020-03-01/index.Rmd
Update slides
- remove infomaps bioregions

---
output:
  xaringan::moon_reader:
    css: ["default-fonts", "default", "custom.css"]
    lib_dir: libs
    mathjax: default
    seal: false
    nature:
      slideNumberFormat: "" # To turn off slide numbers
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["left", "middle", "inverse"]
editor_options:
  chunk_output_type: console
---
class: inverse, middle, title-slide, big-margin
<!-- background-position: 50% 50%
background-image: url(images/title_slide_background.svg)
background-size: contain -->

# 日本のシダ植物多様性の多次元解析
<br>
## 新田ジョエル^1^、海老原淳^2^

^1^スミソニアン国立自然史博物館植物研究部<br>
<span style = 'font-size: 80%;'>https://joelnitta.com</span>

^2^国立科学博物館植物研究部

@@

```{r load-packages, cache = FALSE}
library(drake)
library(knitr)
library(patchwork)
library(scales)
library(scico)
library(magrittr)
library(broom)
library(assertr)
library(sf)
library(tidyverse)
```

@@

# Load targets
loadd(list = c(
  "all_cells",
  "alpha_div_ferns",
  "occ_data_ferns",
  "ja_bioregions_ferns",
  "biodiv_results_pteridos",
  "species_motifs_ferns_6L",
  "percent_sex_dip_ferns"), 
  cache = ja_fern_cache)

@@

# Fern richness
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "richness", 
  metric_title = "No. species ",
  label = "Richness") +
  scale_fill_scico(palette = "bamako", na.value="grey") +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    limits = c(24, 46)) +
  theme(
    plot.subtitle = element_blank())

@@

@@

## Highest bioregional diversity in S islands

```{r mapequation-prep, include = FALSE}
# Convert all_cells dataframe to sf object
all_cells_sf <-
  purrr::map2(all_cells$longitude, all_cells$latitude, ~st_point(c(.x, .y))) %>%
  st_sfc(crs = 4326) %>%
  st_sf(all_cells, .)

# Join secondary grid cells with bioregions
all_cells_regions <-sf::st_join(
  all_cells_sf,
  ja_bioregions_ferns) %>% 
  select(secondary_grid_code, longitude, latitude, bioregion) %>% 
  as_tibble %>%
  mutate(bioregion = as.factor(bioregion)) %>%
  filter(secondary_grid_code %in% occ_data_ferns$secondary_grid_code)
```

```{r mapequation-plot, out.width = "120%", fig.asp = 0.7}
ggplot(all_cells_regions, aes(x = longitude, y = latitude, fill = bioregion)) +
  geom_tile() +
  coord_quickmap(
    xlim = c(pull(occ_data_ferns, longitude) %>% min %>% floor,
             pull(occ_data_ferns, longitude) %>% max %>% ceiling),
    ylim = c(pull(occ_data_ferns, latitude) %>% min %>% floor,
             pull(occ_data_ferns, latitude) %>% max %>% ceiling)
  ) +
  scale_fill_brewer(palette = "Set1") +
    scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
    scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  labs(fill = "Bioregion") +
  theme(
    panel.background = element_blank(),
      panel.grid.major = element_line(size = 0.1, color = "dark grey"),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
    axis.ticks = element_blank()
  )

## Honshu has significant ecological structure

```{r ecostructure}
# For easy manipulation of plot size
tweak = 0.5

# Output pie plot without basemap
ecos_plot_pie2(
  omega =  species_motifs_ferns_6L$omega,
  image_width = 6000*tweak,
  image_height = 4800*tweak,
  radius = 0.05,
  no_map = TRUE,
  blank_bg = "white",
  long_lim = c(122, 146),
  lat_lim = c(24, 46),
  path = here::here("presentations/jsps_2020-03-01/images/species_motifs_ferns_6L.png")
)

# crop image, save as species_motifs_ferns_6L_cropped.png
```

.center[<img src="images/species_motifs_ferns_6L_cropped.png" alt="drawing" height="600">]

@@

```{r canape-plot, out.width = "120%", fig.asp = 0.7}
ggplot(biodiv_results_pteridos, aes(x = longitude, y = latitude, fill = canape)) +
  geom_tile() +
  coord_quickmap(
    xlim = c(pull(occ_data_ferns, longitude) %>% min %>% floor,
             pull(occ_data_ferns, longitude) %>% max %>% ceiling),
    ylim = c(pull(occ_data_ferns, latitude) %>% min %>% floor,
             pull(occ_data_ferns, latitude) %>% max %>% ceiling)
  ) +
  scale_fill_brewer(palette = "Set1", na.value = "grey80") +
    scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
    scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  labs(fill = "Endemism") +
  theme(
    panel.background = element_blank(),
      panel.grid.major = element_line(size = 0.1, color = "dark grey"),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
    axis.ticks = element_blank()
  )
```

@@

```{r canape-sex, out.width = "120%", fig.asp = 0.7}

@@

canape_sex_mode <-
  biodiv_results_pteridos %>%
  mutate(latitude = round(latitude, 3)) %>%
  mutate(longitude = round(longitude, 3)) %>%
  inner_join(
    percent_sex_dip_ferns_round,
    by = c("latitude", "longitude")
  ) %>%
  assert(is_uniq, secondary_grid_code) %>%
  filter(!is.na(canape))

@@

ggplot(canape_sex_mode, aes(x = canape, y = percent_sex_dip)) +
  geom_jitter() +
  geom_boxplot() +
  coord_flip()

@@

>>>>

---
output:
  xaringan::moon_reader:
    css: ["default-fonts", "default", "custom.css"]
    lib_dir: libs
    mathjax: default
    seal: false
    nature:
      slideNumberFormat: "" # To turn off slide numbers
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      titleSlideClass: ["left", "middle", "inverse"]
editor_options:
  chunk_output_type: console
---
class: inverse, middle, title-slide, big-margin
background-color: #000000
background-position: 50% 50%
background-image: url(images/title_background.png)
background-size: contain

# 日本のシダ植物多様性の<br>多次元解析
<br>
## 新田ジョエル<sup>1</sup>、海老原淳<sup>2</sup>

<sup>1</sup>スミソニアン国立自然史博物館・植物研究部<br>
<span style = 'font-size: 80%;'>https://joelnitta.com</span>

<sup>2</sup>国立科学博物館・植物研究部

@@

```{r load-packages, cache = FALSE}
library(magick)
library(gt)
library(drake)
library(knitr)
library(patchwork)
library(scales)
library(scico)
library(magrittr)
library(broom)
library(assertr)
library(sf)
library(tidyverse)
```

@@

# Load targets
loadd(list = c(
  "taxon_id_map",
  "all_cells",
  "alpha_div_ferns",
  "occ_data_ferns",
  "ja_bioregions_ferns",
  "biodiv_results_ferns",
  "species_motifs_ferns_6L",
  "percent_sex_dip_ferns"), 
  cache = ja_fern_cache)

@@

# Fern richness
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "richness", 
  metric_title = "No. species ",
  label = "Richness") +
  scale_fill_viridis_c() +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    limits = c(24, 46)) +
  theme(
    plot.subtitle = element_blank())

@@

```{r title-slide-image}
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "richness", 
  metric_title = "No. species ",
  label = "Richness") +
  scale_fill_viridis_c() +
  theme_void() +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    legend.position = "none", 
     panel.background = element_rect(fill = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )

ggsave(plot = a, here::here("presentations/jsps_2020-03-01/images/title_slide_part_a.png"))

b <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "phy_mpd_obs_z", 
  metric_title = "MPD",
  label = "Mean Phylogenetic Distance") +
  scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3)
    )) +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    limits = c(24, 46)) +
  theme_void() +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    legend.position = "none", 
     panel.background = element_rect(fill = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )

ggsave(plot = b, here::here("presentations/jsps_2020-03-01/images/title_slide_part_b.png"))

@@

@@

```{r canape-plot, out.width = "120%", fig.asp = 0.7}
canape_cols <- RColorBrewer::brewer.pal(4, "Dark2") %>%
  set_names(c("mixed", "neo", "palaeo", "super"))

ggplot(biodiv_results_ferns, aes(x = longitude, y = latitude, fill = canape)) +
  geom_tile() +
  coord_quickmap(
    xlim = c(pull(occ_data_ferns, longitude) %>% min %>% floor,
             pull(occ_data_ferns, longitude) %>% max %>% ceiling),
    ylim = c(pull(occ_data_ferns, latitude) %>% min %>% floor,
             pull(occ_data_ferns, latitude) %>% max %>% ceiling)
  ) +
  scale_fill_manual(values = canape_cols, na.value = "grey80") +
    scale_y_continuous(labels = scales::label_number(suffix = "\u00b0N", accuracy = 1)) +
    scale_x_continuous(labels = scales::label_number(suffix = "\u00b0E", accuracy = 1)) +
  labs(fill = "Endemism") +
  theme(
    panel.background = element_blank(),
      panel.grid.major = element_line(size = 0.1, color = "dark grey"),
      panel.grid.minor = element_blank(),
      axis.title = element_blank(),
    axis.ticks = element_blank()
  )
```

@@

<br>

.center[
```{r canape-sex, fig.asp = 0.7}

@@

canape_sex_mode <-
  biodiv_results_ferns %>%
  mutate(latitude = round(latitude, 3)) %>%
  mutate(longitude = round(longitude, 3)) %>%
  inner_join(
    percent_sex_dip_ferns_round,
    by = c("latitude", "longitude")
  ) %>%
  assert(is_uniq, secondary_grid_code) %>%
  filter(!is.na(canape))

@@

canape_sex_mode %>%
  group_by(canape) %>%
  summarize(
    mean_sex_dip = mean(percent_sex_dip, na.rm = TRUE),
    sd_sex_dip = sd(percent_sex_dip, na.rm = TRUE)
  ) %>%
  mutate(
    upper = mean_sex_dip + sd_sex_dip,
    lower = mean_sex_dip - sd_sex_dip
  ) %>%
  mutate(canape = fct_reorder(canape, mean_sex_dip)) %>%
  ggplot(aes(x = canape, y = mean_sex_dip, fill = canape)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = canape_cols, na.value = "grey80", guide = FALSE) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.25) +
  labs(
    x = "Type of endemicity",
    y = "% sexual diploids"
  ) +
  jntools::standard_theme()

```
]

---
class: inverse, middle
background-color: #000000

## Ecological structure of Japanese ferns, *K* = 6

```{r ecos-map-prep}
# Write out ecostructure plots using modified version of ecos_plot_pie(),
# then crop them to the right size.
# (ggscatterpie is too slow, can't map aesthetics)

# Set a base tweaking factor for easy manipulation of plot size
tweak = 0.5

# Use color brewer "Set1" for the 6 motifs (k=6)
ecos_cols <- RColorBrewer::brewer.pal(6, "Set1")

# Define a function to dim all other colors (using alpha)
# besides the focus color
apply_focus <- function (cols, focus, alpha = 0.15) {
  new_cols <- scales::alpha(cols, alpha)
  new_cols[[focus]] <- cols[[focus]]
  new_cols
}

# Make list of colors to use for writing out plots
cols_list <- purrr::map(1:6, ~apply_focus(ecos_cols, .)) %>%
  c(list(ecos_cols), .) %>%
  set_names(c("all", paste0("focus_", 1:6)))

# Loop over outputting pie plot without basemap
purrr::walk2(
  cols_list,
  names(cols_list),
  ~ecos_plot_pie2(
    omega =  species_motifs_ferns_6L$omega,
    color = .x,
    image_width = 6000*tweak,
    image_height = 4800*tweak,
    radius = 0.05,
    no_map = TRUE,
    blank_bg = "black",
    long_lim = c(122, 146),
    lat_lim = c(24, 46),
    path = here::here(
      glue::glue("presentations/jsps_2020-03-01/images/species_motifs_ferns_{.y}.png")
    )
  )
)

# Define a function to crop ecostructure map plot
crop_ecos_plot <- function (image_path) {

  image_read(image_path) %>%
    # Crops to 2410x2410, starting 525 to the right and 150 down from UL corner
    image_crop("2140x2140+525+150") %>%
    image_write(
      path = image_path %>% 
        fs::path_ext_remove() %>%
        paste0("_crop.png")
    )
}

# Loop over plots and crop them
here::here(glue::glue("presentations/jsps_2020-03-01/images/species_motifs_ferns_{names(cols_list)}.png")) %>%
  walk(crop_ecos_plot)
```

```{r ecos-species-prep}
# Make overall dataframe of species contributions to motifs based on theta
species_ranking <-
species_motifs_ferns_6L$theta %>%
  as.data.frame %>%
  rownames_to_column("taxon_id") %>%
  as_tibble %>%
  pivot_longer(-taxon_id, names_to = "motif", values_to = "score") %>%
  left_join(taxon_id_map, by = "taxon_id") %>% 
  select(motif, taxon, score) %>%
  mutate(taxon = str_replace_all(taxon, "_", " ")) %>%
  mutate(score = round(score, 2))

# Define function for making species ranking table
make_species_ranking_table <- function (df, motif_select, slice_n = 15) {

  df %>%
  filter(motif == motif_select) %>%
  arrange(desc(score), taxon) %>% 
  slice(1:slice_n) %>%
  select(taxon, score) %>%
  gt() %>%
  tab_options(
    row.striping.include_table_body = FALSE,
    row.striping.include_stub = FALSE,
    table.background.color = "#000000",
    column_labels.background.color = "#000000") %>%
  tab_style(
    style = list(
      cell_fill(color = "#000000"),
      cell_text(color = "white", size = "medium")
      ),
    locations = cells_column_labels(columns = everything())
    ) %>%
  tab_style(
    style = list(
      cell_fill(color = "#000000"),
      cell_text(color = "white", size = "small")
      ),
    locations = cells_data(
      columns = everything())
  ) %>%
  tab_style(
    style = list(
      cell_text(style = "italic")
      ),
    locations = cells_data(
      columns = vars(taxon))
  )
}

@@

.center[<img src="images/species_motifs_ferns_all_crop.png" alt="drawing" height = "600">]

---
class: inverse, middle
background-color: #000000

.pull-left-wide[<img src="images/species_motifs_ferns_focus_2_crop.png" alt="drawing">]

.pull-right-narrow[
```{r ecos-ranking-2}
make_species_ranking_table(species_ranking, 2)
```
]

---
class: inverse, middle
background-color: #000000

.pull-left-wide[<img src="images/species_motifs_ferns_focus_3_crop.png" alt="drawing">]

.pull-right-narrow[
```{r ecos-ranking-3}
make_species_ranking_table(species_ranking, 3)
```
]

---
class: inverse, middle
background-color: #000000

.pull-left-wide[<img src="images/species_motifs_ferns_focus_4_crop.png" alt="drawing">]

.pull-right-narrow[
```{r ecos-ranking-4}
make_species_ranking_table(species_ranking, 4)
```
]

---
class: inverse, middle
background-color: #000000

.pull-left-wide[<img src="images/species_motifs_ferns_focus_5_crop.png" alt="drawing">]

.pull-right-narrow[
```{r ecos-ranking-5}
make_species_ranking_table(species_ranking, 5)
```
]

---
class: inverse, middle
background-color: #000000

.pull-left-wide[<img src="images/species_motifs_ferns_focus_6_crop.png" alt="drawing">]

.pull-right-narrow[
```{r ecos-ranking-6}
make_species_ranking_table(species_ranking, 6)
```
]

---
class: inverse, middle
background-color: #000000

.pull-left-wide[<img src="images/species_motifs_ferns_focus_1_crop.png" alt="drawing">]

.pull-right-narrow[
```{r ecos-ranking-1}
make_species_ranking_table(species_ranking, 1)
```
]
